---
title: "General report on sequences and clusters"
format: html
editor: visual

params:
  length: "null"
  length_network: "null"
  network: "null"
---

```{r}
#| echo: false
#| output: false
#| warning: false

library(dplyr)
library(ggplot2)
library(knitr)
```

# Sequence length

```{r}
#| echo: false
#| warning: false

length <- read.table(params$length, header = FALSE, sep = "\t", col.names = c("protein_accession", "length"))
length.network <- read.table(params$length_network, header = FALSE, sep = "\t", col.names = c("protein_accession", "length", "inflation"))
```

## Before clustering

Sequence lengths before construction of the sequence similarity network and clustering with the Markov CLustering algorithm.

```{r}
#| echo: false
#| warning: false

plt.length <- length |> 
   ggplot(aes(x = length)) +
   geom_histogram(fill = "steelblue", color = "black") +
   theme_linedraw() +
   labs(x = "Squence length",
        y = "Number of sequences",
        caption = "results/sequences/sequence_length.tsv") +
   theme(plot.caption = element_text(face = "italic", size = 11))
print(plt.length)
```

## After clustering

Sequence length distribution after construction of the sequence similarity network and clustering with the Marov CLustering algorithm.

```{r}
#| echo: false
#| warning: false

plt.length.network <- length.network |> 
   ggplot(aes(x = length, fill = inflation)) +
   geom_histogram(color = "black") +
   scale_fill_brewer(palette="Spectral") +
   facet_wrap(~ inflation, ncol=3, labeller = labeller(inflation = ~ gsub("_", " ", .x))) +
   theme_linedraw() +
   labs(x = "Squence length",
        y = "Number of sequences",
        caption = "results/sequences/sequence_length_network.tsv") +
   theme(legend.position = "none",
         strip.background = element_rect(fill = "#6a9be3"),
         strip.text = element_text(colour = "white", size = 12, face = "bold"),
         plot.caption = element_text(face = "italic", size = 11))

print(plt.length.network)
```

## Summary

Summary of sequences (minimum, maximum, average and median length), as well as the number of sequences before running LAGOON-MCL and after clustering.

```{r}
#| echo: false
#| warning: false

stats.length <- length |> 
   summarise(inflation = "before clustering",
             count.seq = n(),
             min.length = min(length),
             max.length = max(length),
             mean.length = mean(length),
             median.length = median(length))

stats.length.network <- length.network |> 
   group_by(inflation) |> 
   summarise(count.seq = n(),
             min.length = min(length),
             max.length = max(length),
             mean.length = mean(length),
             median.length = median(length))

stats.all <- rbind(stats.length, stats.length.network)
stats.all$inflation <- gsub("_", " ", stats.all$inflation)
colnames(stats.all) <- c("", "No. of sequences", "Minimum length", "Maximum length", "Average length", "Median length")

kable(stats.all)
```

# Clusters

```{r}
#| echo: false
#| warning: false

network <- read.table(params$network, header = FALSE, sep = "\t", col.names = c("cluster_id", "protein_accession", "inflation"))
```

## Distribution

Cluster size distribution for different inflations parameters.

```{r}
#| echo: false
#| warning: false

plt.network <- network |> 
   group_by(inflation, cluster_id) |> 
   summarise(count = n()) |> 
   ggplot(aes(x = count, fill = inflation)) +
   geom_histogram(binwidth = 1, color = "black") +
   scale_fill_brewer(palette="Spectral") +
   facet_wrap(~ inflation, ncol=3, labeller = labeller(inflation = ~ gsub("_", " ", .x))) +
   theme_linedraw() +
   labs(x = "Cluster size",
        y = "Number of clusters",
        caption = "results/") +
   theme(legend.position = "none",
         strip.background = element_rect(fill = "#6a9be3"),
         strip.text = element_text(colour = "white", size = 12, face = "bold"),
         plot.caption = element_text(face = "italic", size = 11))

print(plt.network)
```

## Summary

Cluster summary (minimum, maximum, mean and median cluster size) and number of clusters per inflation parameter

```{r}
#| echo: false
#| warning: false

stats.clst <- network |> 
   group_by(inflation, cluster_id) |> 
   summarise(count = n()) |> 
   summarise(count.clst = sum(count),
             min.clst = min(count),
             max.clst = max(count),
             mean.clst = mean(count),
             median.clst = median(count))

stats.clst$inflation <- gsub("_", " ", stats.clst$inflation)
colnames(stats.clst) <- c("Inflation", "No. of cluster", "Minimum size", "Maximum size", "Average size", "Median size")

kable(stats.clst)
```
